---
title: "ESM 206 Lab 5"
author: "Madeline Oliver"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 Objectives

- Learn to parse dates with 'lubridate'
- Get counts of observations with count()
- Uncount() rows
- One and two-sample t-test
- Create a heatmap with geom_title()


```{r, include = FALSE}
# Read in data and attach packages

library(tidyverse)
library(here)
library(janitor)

lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"),
                              na = "-99999") %>%
  janitor::clean_names()


```

Use tidyr::uncount() function to convert our lobster data from frequency format to case format

```{r}

lobster_tidy <- lobster_abundance %>%
  tidyr::uncount(lobster_count)

```

#### Exploratory data visualization

We're only going to consider 'site' as our variable of interest

```{r}

ggplot(lobster_tidy, aes(x = site, 
                         y = size_mm)) +
  geom_jitter(aes(color = site),
              alpha = 0.5,
              width = 0.2)

# Histogram

ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

ggplot(lobster_tidy, aes(sample = size_mm)) + 
  geom_qq() +
  facet_wrap(~site)

```


#### Convert the 'date' column to class 'Date'

We'll use the 'lubridate' package to convert to Date format, and then to help us easily parse month and year

```{r}

lobster_date <- lobster_tidy %>%
  mutate(
    date_new = lubridate::mdy(date)
  )

```

Now let's parse year and month using lubridate::month() and lubridate::year()

```{r}

lobster_parse_date <- lobster_date %>%
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )

```

Now let's find counts of observed lobsters based on different hierarchical groupings:

First: let's count lobsters by year and month:

```{r}
lobster_ym <- lobster_parse_date %>%
  dplyr::count(obs_year, obs_month)

lobster_y <- lobster_parse_date %>%
  dplyr::count(obs_year)

lobster_site <- lobster_parse_date %>%
  dplyr::count(site)

```

If we want to create a summary table that contains statistics OTHER than counts by group, it's easier to use group_by() + n()

```{r}

lobster_summary <- lobster_parse_date %>%
  group_by(site) %>%
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm= TRUE),
    lobster_number = n()
  )

```

#### Find confidence intervals

Use t.test() function to find confidence intervals for on sample and perform t-test to compare means of two samples (... this will be covered conceptually in lectures Week 6)

```{r}

ivee_lobster <- lobster_tidy %>%
  filter(site == "IVEE") %>%
  pull(size_mm)

t.test(ivee_lobster)

```

#### Two-sample t-test to compare means

We want to ask: Is there a significant difference in lobster lengths at Napas and Mohawk reefs?

We've done our nevessary exploratory analyses to determine that a 2-sample t-test for means comparison is appropriate.

```{r}

napl_sample <- lobster_tidy %>%
  filter(site == "NAPL") %>%
  pull(size_mm)

mohk_sample <- lobster_tidy %>%
  filter(site == "MOHK") %>%
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, napl_sample)
mn_ttest

```

There is a significant difference in lobster lengths between Naples and Mohawk Reef (t(df) = statistic, p< 0.001, alpha = 0.05).

```{r}
lobster_mn <- lobster_tidy %>%
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <- t.test(size_mm ~ site, data = lobster_mn)
mn_ttest2

mn_ttest2$p.value
mn_ttest2$statistic

```

Here is the p-value: `r mn_ttest2$p.value`

#### Now: a heatmap!

```{r}
lobster_ys <- lobster_parse_date %>%
  count(obs_year, site)

ggplot(data = lobster_ys, 
       aes(x = obs_year,
           y = site)) +
  geom_tile(aes(fill = n))

```





